Study 2 Analysis

### 2024-11-04 misc/indiv.qmd
update
s1 discrepancy checks; index page reset

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",value<20000) |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(state~matched_goal,scales="free_y")
```

>>>>

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",state=="CAL",value<20000) |> 
  ggplot(aes(x=appliance,y=value)) + 
  geom_jitter(aes(color=matched_goal)) +
geom_point(aes(y=state_avg),color="red",size=10) +
  geom_point(aes(y=family),color="blue",size=5)

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",value<20000,family<25000) |> 
  ggplot(aes(x=appliance,y=value)) + 
  geom_jitter(aes(color=matched_goal)) +
 # gghalves::geom_half_violin(aes(fill=matched_goal)) +
  geom_boxplot(aes(fill=matched_goal),alpha=.5) +
geom_point(aes(y=state_avg),color="red",size=7,shape=8) +
  geom_point(aes(y=family),color="blue",size=7,shape=17) +
  facet_wrap(~state)

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",value<20000,family<25000) |> 
  ggplot(aes(x=appliance,y=state_p_dif)) + 
  geom_jitter(aes(color=matched_goal)) +
  geom_point(aes(y=state_avg),color="red",size=10) +
  geom_point(aes(y=family),color="blue",size=5)

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal,appliance) |> 
summarize(total_kWh = sum(value),orig_kWh=sum(family), 
            adjust=orig_kWh-total_kWh,
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL") |> 
  ggplot(aes(x=appliance,y=adjust)) + 
  geom_jitter(aes(color=matched_goal)) +
  facet_wrap(~state)

### 2024-11-04 misc/outlier.qmd
indiv patterns page
s1 discrepancy checks; index page reset

@@

>>>>

 s1 |> filter(state_avg==29046)
# A tibble: 2 × 21
# Groups:   id [1]
     id state1   state2 refClass calc  edu   modality state id2   Duration__in_seconds_ appliance family state_avg plan  value block
  <int> <chr>    <chr>  <chr>    <chr> <fct> <chr>    <chr> <chr>                 <dbl> <chr>      <dbl> <dbl+lbl> <fct> <dbl> <dbl>
1    12 Colorado Texas  kWh      Used… Grad… A        TEX   R_ab…                   644 Total kWh  29046 29046     plan1 29046     2
2    12 Colorado Texas  kWh      Used… Grad… A        TEX   R_ab…                   644 Total kWh  29046 29046     plan2 29046     2
ℹ 5 more variables: pct_goal <dbl>, change <dbl>, state_p_dif <dbl>, state_f_dif <dbl>, less_avg <lgl>

@@

state_code_map <- c(
  "CAL" = "California",
  "TEX" = "Texas",
  "COL" = "Colorado",
  "MASS" = "Massachusetts"
)

### 2024-11-04 study1.qmd
indiv patterns page
s1 discrepancy checks; index page reset

s1 |> 
    group_by(state, appliance, refClass) |> 
    mutate(n_distinct_family = n_distinct(family),
           n_distinct_state_avg = n_distinct(state_avg), 
           distinct_family = paste(unique(family), collapse = ","),
           distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
    filter(n_distinct_family > 1 | n_distinct_state_avg > 1) |> 
    group_by(state, appliance, refClass, family, state_avg) |> 
    mutate(state_avg_count = n_distinct(state_avg),
           family_count = n()) |> 
    ungroup() |> 
    group_by(state, appliance, refClass, state_avg) |> 
    mutate(state_avg_count = n()) |> 
    distinct(state, appliance, refClass, family, state_avg, family_count, state_avg_count) |> 
  arrange(state,appliance,refClass)

@@

>>>>

s1 |> 
    group_by(state, appliance, refClass) |> 
    mutate(n_distinct_family = n_distinct(family),
           n_distinct_state_avg = n_distinct(state_avg), 
           distinct_family = paste(unique(family), collapse = ","),
           distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
    filter(n_distinct_family > 1 | n_distinct_state_avg > 1,plan=="plan1") |> 
    group_by(state, appliance, refClass, family, state_avg) |> 
    mutate(state_avg_count = n_distinct(state_avg),
           family_count = n()) |> 
    ungroup() |> 
    group_by(state, appliance, refClass, state_avg) |> 
    mutate(state_avg_count = n()) |> 
    distinct(state, appliance, refClass, family, state_avg, family_count, state_avg_count) |> 
  arrange(state,appliance,refClass)

@@

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |>
  filter(ResponseId == "R_vB56XspT1LqhrkB") |>
  # Select columns with non-NA values and not empty strings, handling character and numeric columns separately
  select(
    where(~ is.numeric(.) && !all(is.na(.))),
    where(~ is.character(.) && !all(is.na(.)) && !all(. == ""))
  ) |>
  # Remove columns that start with "AIN"
  select(-starts_with("AIN"))

### 2024-11-04 study2.qmd
outlier checks and some model updates
indiv patterns page
s1 discrepancy checks; index page reset

library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)

@@

s2 |> 
  mutate(
    rounded = if_else(ROUNDED == 1, "Not Rounded", "Rounded"),
    calc = if_else(calc == 0, "No Calculator", "Used Calculator")
  ) |> 
  count(calc, refClass, rounded) |>
  pivot_wider(
    names_from = c(calc, rounded),
    values_from = n,
    values_fill = 0
  ) |>
  gt() |>
  tab_header(
    title = md("**Count of Items by Calculator Use, Reference Class, and Rounding**")
  ) |>
  cols_align(align = "left") |>
  tab_spanner_delim(delim = "_") |>
  cols_label(
    `No Calculator_Not Rounded` = "Not Rounded",
    `No Calculator_Rounded` = "Rounded",
    `Used Calculator_Not Rounded` = "Not Rounded",
    `Used Calculator_Rounded` = "Rounded"
  )

>>>>

library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable,ggh4x)

@@

# s2 |> 
#   mutate(
#     rounded = if_else(ROUNDED == 1, "Not Rounded", "Rounded"),
#     calc = if_else(calc == 0, "No Calculator", "Used Calculator")
#   ) |> 
#   count(calc, refClass, rounded) |>
#   pivot_wider(
#     names_from = c(calc, rounded),
#     values_from = n,
#     values_fill = 0
#   ) |>
#   gt() |>
#   tab_header(
#     title = md("**Count of Items by Calculator Use, Reference Class, and Rounding**")
#   ) |>
#   cols_align(align = "left") |>
#   tab_spanner_delim(delim = "_") |>
#   cols_label(
#     `No Calculator_Not Rounded` = "Not Rounded",
#     `No Calculator_Rounded` = "Rounded",
#     `Used Calculator_Not Rounded` = "Not Rounded",
#     `Used Calculator_Rounded` = "Rounded"
#   )
